---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
library(tidyverse)

```


```{css}
body {
  font-family: 'Times New Roman', sans-serif;
  background-color: #f5f5f5;
}

h2 {
  color: #2c3e50;
  border-bottom: 3px solid #3498db;
  padding-bottom: 10px;
  margin-top: 30px;
}

h3 {
  color: #34495e;
  margin-top: 25px;
}
```

## Auckland Public Transport: Commuter Observations Analysis
This dynamic report presents an analysis of public transport observations collected over a two-week period on a fixed route between home and university. The data captures key aspects of service quality including punctuality, crowding levels, traffic congestion, and travel duration.

```{r}

# Read and prepare data
logged_data <- read_csv("Auckland Public Transport Observation.csv")

# Change the name
latest_data <-logged_data %>%
  rename(
    timestamp = 1,
    date = 2,
    time = 3,
    direction = 4,
    on_time = 5,
    fullness = 6,
    congestion = 7,
    passenger_groups = 8,
    duration = 9
  )

glimpse(latest_data)

```
## Key Summary Statistics

```{r}

avg_duration_home_to_uni <- latest_data %>%
  filter(direction == "Home → University") %>%
  summarise(avg = mean(duration, na.rm = TRUE))

paste0("The average travel time from Home to University was ", round(avg_duration_home_to_uni$avg, 1), " minutes.")


avg_duration_uni_to_home <- latest_data %>%
  filter(direction == "University → Home") %>%
  summarise(avg = mean(duration, na.rm = TRUE))

paste0("The average travel time from University to Home was ", round(avg_duration_uni_to_home$avg, 1), " minutes.")

total_trips <- nrow(latest_data)

paste0("A total of ", total_trips, " trips were recorded during the observation period.")

on_time_count <- latest_data %>%
  filter(on_time == "On time") %>%
  nrow()

paste0("Out of ", total_trips, " trips, ", on_time_count, " services arrived on time.")

on_time_percentage <- (on_time_count / total_trips) * 100

paste0("Overall, ", round(on_time_percentage, 1), "% of services were punctual.")

congestion_count <- latest_data %>%
  filter(congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()

paste0("Traffic congestion was observed in ", congestion_count, " out of ", total_trips, " trips.")
```

## Service Quality Analysis
### Vehicle Fullness vs On-Time Performance

Understanding how crowding levels affect service punctuality is crucial for commuters. The chart below shows the relationship between vehicle fullness and on-time status.

```{r}

ggplot(data = latest_data) +
  geom_bar(aes(x = fullness, fill = on_time))+
  labs(title = "Vehicle Fullness vs On-Time Performance",
       fill = "On-Time Status",
       caption = "Source: Auckland Public Transport Observation Log",
       x = "Vehicle Fullness Level",
       y = "Number of Observations")

# Calculate percentage of on-time services when very busy
very_busy_on_time <- latest_data %>%
  filter(fullness == "Very Busy", on_time == "On time") %>%
  nrow()

very_busy_total <- latest_data %>%
  filter(fullness == "Very Busy") %>%
  nrow()

very_busy_on_time_pct <- (very_busy_on_time / very_busy_total) * 100

paste0("Even when vehicles were 'Very Busy', ", 
       round(very_busy_on_time_pct, 1), 
       "% of services still arrived on time. This suggests crowding does not necessarily cause delays.")
```

## Traffic Congestion by Travel Direction

Traffic congestion is a major factor affecting bus reliability in Auckland. This chart compares congestion frequency between morning (Home → University) and evening (University → Home) trips.

```{r}

ggplot(data = latest_data) +
  geom_bar(aes(x = congestion, fill = direction))+
  labs(title = "Traffic Congestion Frequency by Travel Direction",
       fill = "Travel Direction",
       caption = "Source: Auckland Public Transport Observation Log",
       x = "Congestion Level",
       y = "Number of Observations")

# Calculate congestion differences by direction
morning_congestion <- latest_data %>%
  filter(direction == "Home → University", congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()

evening_congestion <- latest_data %>%
  filter(direction == "University → Home", congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()

paste0("Traffic congestion was more frequent on evening commutes (University → Home) with ", 
       evening_congestion, 
       " events, compared to ", 
       morning_congestion, 
       " events on morning commutes (Home → University).")
```


