---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```


```{css}
body {
  font-family: 'Times New Roman', sans-serif;
  background-color: #f5f5f5;
}

h2 {
  color: #2c3e50;
  border-bottom: 3px solid #3498db;
  padding-bottom: 10px;
  margin-top: 30px;
}

h3 {
  color: #34495e;
  margin-top: 25px;
}

img {
  display: block;
  margin-left: auto;
  margin-right: auto;
}
```

## Auckland Public Transport: Commuter Observations Analysis
This dynamic report presents an analysis of public transport observations collected over a two-week period on a fixed route between home and university. The data captures key aspects of service quality including punctuality, crowding levels, traffic congestion, and travel duration.

```{r}
library(tidyverse)
library(magick)

# Read and prepare data
logged_data <- read_csv("Auckland Public Transport Observation.csv")

# Change the name
latest_data <-logged_data %>%
  rename(
    timestamp = 1,
    date = 2,
    time = 3,
    direction = 4,
    on_time = 5,
    fullness = 6,
    congestion = 7,
    passenger_groups = 8,
    duration = 9
  )

glimpse(latest_data)

image_read("https://www.nzherald.co.nz/resizer/v2/YLJSXA4TGYJD7VTJPXRBJ2C7HU.jpg?auth=87651fb9a97646e5ed5554fe0f2dcabe92be9e8fdc51950111fa43c0f0a54650&width=1440&height=810&quality=70&smart=true") %>%
  image_scale("800") %>%
  image_border()
```

## Key Summary Statistics
```{r}
avg_duration_home_to_uni <- latest_data %>%
  filter(direction == "Home → University") %>%
  summarise(avg = mean(duration, na.rm = TRUE))


avg_duration_uni_to_home <- latest_data %>%
  filter(direction == "University → Home") %>%
  summarise(avg = mean(duration, na.rm = TRUE))


total_trips <- nrow(latest_data)


on_time_count <- latest_data %>%
  filter(on_time == "On time") %>%
  nrow()

on_time_percentage <- (on_time_count / total_trips) * 100

congestion_count <- latest_data %>%
  filter(congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()

```
::: dynamic-comment

1. The average travel time from Home to University was `r round(avg_duration_home_to_uni$avg, 1)` minutes.

2. The average travel time from University to Home was `r round(avg_duration_uni_to_home$avg, 1)` minutes.

3. A total of `r total_trips` trips were recorded during the observation period.

4. Out of `r total_trips` trips, `r on_time_count` services arrived on time.

5. Overall, `r round(on_time_percentage, 1)`% of services were punctual.

6. Traffic congestion was observed in `r congestion_count` out of `r total_trips` trips.

:::

## Service Quality Analysis
### Vehicle Fullness vs On-Time Performance

Understanding how crowding levels affect service punctuality is crucial for commuters. The chart below shows the relationship between vehicle fullness and on-time status.

```{r}
ggplot(data = latest_data) +
  geom_bar(aes(x = fullness, fill = on_time))+
  labs(title = "Vehicle Fullness vs On-Time Performance",
       fill = "On-Time Status",
       caption = "Source: Auckland Public Transport Observation Log",
       x = "Vehicle Fullness Level",
       y = "Number of Observations")

# Calculate percentage of on-time services when very busy
very_busy_on_time <- latest_data %>%
  filter(fullness == "Very Busy", on_time == "On time") %>%
  nrow()

very_busy_total <- latest_data %>%
  filter(fullness == "Very Busy") %>%
  nrow()

very_busy_on_time_pct <- (very_busy_on_time / very_busy_total) * 100

image_read("https://cities-today.com/wp-content/uploads/2021/07/Auckland-City-Council-e1626183282122.jpg") %>%
  image_scale("800") %>%
  image_border()

```

::: dynamic-comment

Even when vehicles were "Very Busy", `r round(very_busy_on_time_pct, 1)`% of services still arrived on time. This suggests crowding does not necessarily cause delays.

:::
  
       
## Traffic Congestion by Travel Direction

Traffic congestion is a major factor affecting bus reliability in Auckland. This chart compares congestion frequency between morning (Home → University) and evening (University → Home) trips.

```{r}
ggplot(data = latest_data) +
  geom_bar(aes(x = congestion, fill = direction))+
  labs(title = "Traffic Congestion Frequency by Travel Direction",
       fill = "Travel Direction",
       caption = "Source: Auckland Public Transport Observation Log",
       x = "Congestion Level",
       y = "Number of Observations")

# Calculate congestion differences by direction
morning_congestion <- latest_data %>%
  filter(direction == "Home → University", congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()

evening_congestion <- latest_data %>%
  filter(direction == "University → Home", congestion != "No traffic congestion", congestion != "Not applicable") %>%
  nrow()


image_read("https://i0.wp.com/newsroom.co.nz/wp-content/uploads/2026/03/congestion-on-auckland-motorway-1.jpg?fit=1024%2C677&ssl=1") %>%
  image_scale("800") %>%
  image_border()

```

::: dynamic-comment

Traffic congestion was more frequent on evening commutes (University → Home) with `r evening_congestion` events, compared to `r morning_congestion` events on morning commutes (Home → University).
::: 